### 2015-06-26 text/ms.Rmd
Major editing of paper, especially discussion

Here, we develop ensemble models to estimate the population status and trajectory of exploited fish populations. We explore a variety of ensemble approaches applied to both simulated and real-world fish stocks and compare predictive performance. In developing these models we highlight the types of challenges one encounters when forming and evaluating ensemble models. We find that ensembles generally provide improved predictive performance than any single model---especially when considered across multiple dimensions of predictive ability and across multiple datasets. Furthermore, the output from ensemble models can provide better mechanistic understanding of how individual models of status and trajectory perform.

@@

Ensemble models, and in particular the machine learning ensemble models (random forest and GBM), generally improved estimates of status and trajectory (Fig. \ref{hexagon}e--h, Fig. \ref{scatter-sim-slope}e--h). Compared to the individual models, machine learning ensembles improved accuracy (median absolute proportional error) by `r mean_sim$mach_vs_ind_mare_fold`%, increased rank-order correlation from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and reduced bias (median proportional error) from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range` (Figs \ref{performance}a). These ensembles also had better ability to distinguish if simulated stocks were above or below $B/B_\mathrm{MSY} = 1$ (Fig. \ref{roc-sim}). Results were similar when predicting trajectory: compared to individual models, machine learning ensembles improved accuracy by `r slope_sim$mach_vs_ind_mare_fold`%, increased rank-order correlation from `r slope_sim$ind_corr_range` to `r slope_sim$mach_corr_range`, and reduced bias from `r slope_sim$ind_mre_range` to `r slope_sim$mach_mre_range` (Fig. \ref{performance-sim-slope}). The ensemble models that simply took a mean of the individual models ranked slightly behind the best individual model for estimating status (CMSY; Fig. \ref{performance}a) and had slightly lower correlation but higher accuracy than the best individual model at predicting slope (SSCOM; Fig. \ref{performance-sim-slope}).

@@

Ensemble models provide a useful approach to situations where conservation and management decisions must be made on the basis of multiple, potentially contrasting estimates of status---a situation that is common to many ecological settings beyond fisheries. Compared to individual models of fish population status, ensemble models were consistently the best or among the best across three performance dimensions, two response variables, two datasets, and multiple ensemble model types. Our results suggest choosing an ensemble model that allows for non-parametric relationships, such as machine learning methods---these methods provide added insight into individual model behaviour and generally performed the best---but even a simple average of predictions across multiple models may be useful.

Certain conditions will make some ensemble models more effective than others. First, ensemble models will be most effective when they are comprised of diverse individual models that make different assumptions about underlying dynamics, or make contrasting but plausible assumptions about parameter values [e.g. @ali1996; @tebaldi2007]. Such individual models would be expected to perform well across a variety of conditions and an ensemble model can exploit the best predictive performance of each. Second, ensemble models will be most effective when they are not overfit to the training dataset. The probability of overfitting ensemble models generally increases with the number of individual models [e.g. @caruana2004]. Cross-validation testing, information theoretic methods (e.g. AIC), and methods that are robust to over-fitting such as random forests (REF), may help avoid overfitting ensemble models. We note that our simplest ensemble model, a mean of the individual model predictions, performed approximately as well as the complex machine learning models when we applied our ensembles to an entirely new dataset (FIG). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest. Cross-validation within a training dataset will provide a optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest.

Ensemble models share similarities with coefficient averaging [@burnham2002] but differ other important ways. Ensemble models and coefficient averaging share the same notion that no one model is correct and multiple models can contribute useful information for inference. A fundamental difference between coefficient averaging and ensemble models is that coefficient averaging focuses on a averaging *coefficients* whereas ensembles instead average *predictions*. By focussing on predictions, ensembles are a more general purpose tool: they do not require information theoretics, they can combine different types of models (e.g. combining parametric and non-parametric models or frequentist and Bayesian predictions), and they can combine predictions in complex non-linear ways that may depend on additional covariates.

A strength of ensemble models is that they can be tailored or specific response variables. For example, we built separate models of mean \bbmsy\\ and the slope of \bbmsy\\. The same set of model weights or non-linear relationships need not hold across different response variables: ensemble models can optimally combine the individual models depending on the response. Ensemble models could be combined to estimate any number of response variables of interest. For example, they could be combined to estimate a binary response such as whether status is above or below some threshold.

@@

* Jim: Other possible applications of ensemble models outside fisheries status and
  trajectory
    * other taxa examples
    * other kinds of response variables

<!--
* Other possible topics:
    * how much better are the ensembles really? translate the results into
      some concrete examples in an absolute not relative sense
    * other types of ensembles: assigning weights, incorporating uncertainty
      around estimates, GAMs, ...
    * when wouldn't you want to use ensembles? perhaps if models represent 
      dichotomous assumptions where either one or the other is right and imply
      different management actions? ...
    * your ideas here...
-->

# References

@@

@@

- @rykiel1996 "Testing ecological models: the meaning of validation" (see for
  performance criteria, theory on model testing and assessment)
-->

>>>>

Here, we develop ensemble models to estimate the population status and trajectory of exploited fish populations. We explore a variety of ensemble approaches applied to both simulated and real-world fish stocks and compare predictive performance. In developing these models we highlight the types of challenges one encounters when forming and evaluating ensemble models. We find that ensembles generally provide improved predictive performance over any single model---especially when considered across multiple dimensions of predictive ability and across multiple datasets. Furthermore, the output from ensemble models can provide better mechanistic understanding of how individual models of status and trajectory perform.

@@

Ensemble models, and in particular the machine learning ensemble models (random
forest and GBM), generally improved estimates of status and trajectory (Fig.
\ref{hexagon}e--h, Fig. \ref{scatter-sim-slope}e--h). Compared to the
individual models, machine learning ensembles improved accuracy (median
absolute proportional error) by 
`r mean_sim$mach_vs_ind_mare_fold`%, 
increased rank-order correlation from `r mean_sim$ind_corr_range` 
to `r mean_sim$mach_corr_range`, 
and reduced bias (median proportional error) from 
`r mean_sim$ind_mre_range` to 
`r mean_sim$mach_mre_range` (Fig. \ref{performance}a). 
These ensembles also had better ability to distinguish if
simulated stocks were above or below $B/B_\mathrm{MSY} = 1$ (Fig.
\ref{roc-sim}). Results were similar when predicting trajectory: compared to
individual models, machine learning ensembles improved accuracy by 
`r slope_sim$mach_vs_ind_mare_fold`%, increased rank-order correlation from 
`r slope_sim$ind_corr_range` to 
`r slope_sim$mach_corr_range`, and reduced bias
from `r slope_sim$ind_mre_range` 
to `r slope_sim$mach_mre_range` (Fig. \ref{performance-sim-slope}). The
ensemble models that simply took a mean of the individual models ranked
slightly behind the best individual model for estimating status (CMSY; Fig.
\ref{performance}a) and had slightly lower correlation but higher accuracy than
the best individual model at predicting slope (SSCOM; Fig.
\ref{performance-sim-slope}).

@@

Ensemble methods provide a useful approach to situations where conservation and
management decisions must be made on the basis of multiple, potentially
contrasting estimates of status---a situation common to many ecological
settings beyond fisheries. Compared to individual models of fish population
status, ensemble models were consistently the best or among the best across
three performance dimensions, two response variables, two datasets, and
multiple ensemble model types. Our results suggest choosing an ensemble model
that allows for non-parametric relationships, such as machine learning
methods---these methods provide added insight into individual model behaviour
and generally performed the best---but even a simple average of predictions
across multiple models may be useful.

Certain conditions will make some ensemble models more effective than others.
First, ensembles will be most effective when they are comprised of diverse
individual models that choose different structural model forms, explore
contrasting but plausible ranges of parameter values, or make errors in
uncorrelated ways [e.g. @ali1996; @dietterich2000; @tebaldi2007]. Such
individual models would be expected to perform well in different conditions and
an ensemble model can exploit the best predictive performance of each. Second,
ensemble models will be most effective when they are not overfit to the
training dataset. Cross-validation testing [e.g. @caruana2004; @hastie2009] and
methods that are robust to over-fitting such as random forests [@breiman2001],
may help avoid overfitting ensemble models. We note that our simplest ensemble
model, an average of individual model predictions, performed approximately as
well as complex machine learning models when we applied our ensembles to an
entirely new dataset (Fig. \ref{performance}b) and averages are a common
ensemble approach in climate science [@ipcc2013]. Third, ensemble
models will be most effective when they are trained on data that are
representative of the dataset of interest [e.g. @knutti2009; @weigel2010].
Cross-validation within a training dataset will provide a optimistically biased
impression of predictive performance if the training dataset fundamentally
differs from the dataset of interest [@hastie2009].

Multi-model inference in the form of coefficient averaging weighted by
information theoretics such as the Akaike Information Criteron (AIC) is a
common analytical approach in ecology [e.g. @burnham2002; @grueber2011]. The
ensemble methods described in this paper share similarities with coefficient
averaging but differ in other important ways. Ensemble models and coefficient
averaging share the long-held notion that multiple working hypotheses can
contribute useful information for inference [@chamberlin1890]. A fundamental
difference is that coefficient averaging focuses on a averaging *coefficients*
whereas ensembles instead average *predictions*. By focussing on predictions,
ensembles do not suffer from the muddled inference of coefficient averaging
when coefficients covary [@cade2015] and are a more general purpose tool: they
do not require information theoretics, they can combine different types of
models (e.g. parametric and non-parametric models or frequentist and
Bayesian predictions), and they can combine predictions in complex non-linear
ways that may depend on additional covariates.

A strength of ensemble models is that they can be tailored or specific response
variables. For example, we built separate ensemble models of mean \bbmsy\\ and
the slope of \bbmsy\\. The same set of model weights or non-linear
relationships need not hold across different response variables. For example,
SSCOM contributed little to the GBM ensemble estimate of status at higher
levels of predicted \bbmsy\\, but contributed strongly to estimates of slope
(TODO FIG). Formally, this fitting of ensemble models to specific quantities of
interest provides an additional calibration step [@rykiel1996]. Individual
models are typically calibrated (i.e. parameters are estimated) based on a
single response variable; however, this model might then provide a biased or
inaccurate estimate of quantities. For example, a data-limited fisheries model
might calibrate parameters by maximizing likelihood with respect to observed
and predicted catches but those parameters may then provide a biased or
inaccurate estimate of the mean and slope of \bbmsy\\. Ensemble methods can
provide a final calibration step to a quantity of interest.

@@

Combining predictions from multiple models via ensemble methods is also broadly
useful in other subfields of ecology and fisheries science. Predictions about
extinction risk are widely used at the national (e.g. the US Endangered Species
Act and the Canadian Species at Risk Act) and international [e.g. the IUCN Red
List, @iucn2015] levels. These risk assessments generally involve fitting
regression models to outcomes for individual species along with predictors of
species risk [e.g. @anderson2011a;@pinsky2011], or population dynamics models to
data for individual species [e.g. @dfo2010]. Both types of models are prone to
error caused by model-misspecification and therefore results are sensitive to
decisions about model structure [@brooks2015]. Although there are options to
account for potential model-misspecification in determination of species risk
[e.g., coefficient averaging, @burnham2002; or semi-parametric methods,
@thorson2014], ensemble methods are a relatively simple way to combine
predictions in a transparent manner. Beyond estimates of status and trajectory,
ensemble methods could be used to increase the robustness of spatial
predictions when designing networks of protected areas [@rassweiler2014], or to
forecast potential spatial shifts in species distribution given climate impacts
[@harsch2014].

<!--
* Other possible topics:
    * how much better are the ensembles really? translate the results into
      some concrete examples in an absolute not relative sense
    * other types of ensembles: assigning weights, incorporating uncertainty
      around estimates, GAMs, ...
    * when wouldn't you want to use ensembles? perhaps if models represent 
      dichotomous assumptions where either one or the other is right and imply
      different management actions? ...
    * your ideas here...

Why do ensembles work?

- 'representational' no one model can contain all 'hypothetical functional
  forms'... many models can cover many hypotheses @dietterich2000
- error cancellation --- especially if multiple models make errors in
  uncorrelated ways [@ali1996]
- they provide an additional step of 'calibration' [@rykiel1996] where the
  predictions are calibrated to data of known or assumed truth about a
  particular aspect of interest in the data (e.g. mean status, slope of status)
  whereas individual model may be calibrated (parameters estimated) with a
  slightly different response in mind (e.g. maximizing likelihood with respect
  to observed and predicted catches)
- we usually do not know which individual model is 'best' [@hagedorn2005];
  ensembles have the property of rarely being far from the best

-->

@@

Why do ensemble methods work?

@@

- @rykiel1996 "Testing ecological models: the meaning of validation" (see for
  performance criteria, theory on model testing and assessment)

model averaging AIC:
f Chamberlin 1890).  multiple woring hypotheses
  (e.g., Johnson and Omland 2004, Hobbs and Hilborn 2006, Burnham et al. 2011,
  52 Grueber et al. 2011). 
-->
